```{r, include=FALSE}
library(ggplot2)
library(dplyr)
theme_custom <- function() {
  
  theme_minimal() %+replace%
    
    theme(
      
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA),
      axis.ticks = element_line(),
      #axis.line.x = element_line(color = "#C5C5C5"),
      axis.line = element_line(color = "black"),
      #axis.ticks = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      #plot.background = element_rect(fill='transparent', color=NA),
      
      plot.title = element_text(             #title
                   #family = font,            #set font family
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = 0,                #left align
                   vjust = 2,
                   color = "black"),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
                   #family = font,            #font family
                   size = 12,
                   hjust = 0,
                   color = "black"),               #font size
      
      plot.caption = element_text(           #caption
                   #family = font,            #font family
                    face = "italic",
                   size = 8,                 #font size
                   hjust = 1,
                   color = "black"),               #right align
      
      axis.title = element_text(             #axis titles
                   #family = font,            #font family
                   size = 12,
                   color = "black"),               #font size
      
      axis.text = element_text(              #axis text
                   #family = font,            #axis famuly
                   size = 12,
                   color = "black"),                #font size
      
      axis.text.x = element_text(            #margin for axis text
                    margin=margin(5, b = 10))
      
    )
}
```

# Topics in Life Insurance

## Computing moments of reserve in timehomogeneous case

In this exercise we consider a time-homogeneous Markov jump process $X$ on the state-space $E=\{1,...,5\}$ defined as such

  1. Active,
  2. Inactive,
  3. Unemployed,
  4. Active after unemployment,
  5. Dead.

The associated intensity matrix $\mathbf \Lambda$ is

$$
\mathbf\Lambda(x)=\mathbf \Lambda=
\begin{bmatrix}
-0.7 & 0.1 & 0.1&    &0.5 \\
 &-0.5 &  &  & 0.5 \\
 & 0.1& -0.7 & 0.1 & 0.5\\
 & 0.1 & & -0.6 & 0.5\\
 &  &  &  & \\
\end{bmatrix}.
$$

Where all zero entries are left blank. We consider the insurance contract that,

  * Insured pays $\rho$ as long as he is alive,
  * Insured recieve 1 while unemployed or inactive and
  * Insured receive 2 if he becomes inactive. If he is active just before the jump this only happends with probability $1/2$.

The contract expires at time $T=1$ and we assume $X(0)=1$. We want to compute the equivalence premium for a constant interest rate $r=0$. 

  a. Determine $\mathbf B(t)$ and $b$.
  b. Decompose $\mathbf \Lambda$ into $\mathbf \Lambda=\mathbf \Lambda^0+\mathbf \Lambda^1$ and compute the rewards $\mathbf R$.
  c. Compute the reserve at time $t=0$ for $\rho = -1$.
  d. Compute the equivalence premium $\hat\rho$ and check that $\mathbf V(0;X(0)=0)=0$.
  e. Compute the contributions $\mathbf C_r^{(2)}$ and use this to compute the second moment $\mathbf V^{(2)}(0;X(0)=0)$.

<details>
<summary>**Solution (a).**</summary>

The matrix $\mathbf B=\left\{b_{ij}\right\}_{i,j\in E}$ is given by $b_{ij}=2$ if $j=2$ and $i\in \{1,3,4\}$ i.e.

$$
\mathbf B=
\begin{bmatrix}
0 & 2 &0 &0 & 0 \\
0 & 0 &0 &0 & 0 \\
0 & 2 &0 &0 & 0 \\
0 & 2 &0 &0 & 0 \\
0 & 0 &0 &0 & 0 \\
\end{bmatrix}.
$$

The continuous payment rate is

$$
b=\begin{pmatrix}
\rho\\
1\\
1\\
\rho\\
0
\end{pmatrix}.
$$

We implement this i R using.

```{r}
rho <- -1
b <- c(rho, 1, 1, rho,0)
B <- t(
  matrix(
    c(0,2,0,0,0,
      0,0,0,0,0,
      0,2,0,0,0,
      0,2,0,0,0,
      0,0,0,0,0), ncol = 5
  )
)
```


</details>

<details>
<summary>**Solution (b).**</summary>

The $\mathbf \Lambda$ is decomposed by setting $(\lambda^1_{12},\lambda^1_{32},\lambda^1_{42})=(\lambda_{12}/2,\lambda_{32},\lambda_{42}/2)$ and the rest to 0 since the payment when transitioning from 1 or 4 only happens with probabiltiy $1/2$. The reward matrix is simply given by $\mathbf R=\mathbf \Lambda^1 \bullet \mathbf B + \Delta b$ i.e.

$$
\mathbf R = 
\begin{bmatrix}
\rho & 2\lambda_{12}^1 &0 &0 & 0 \\
0 & 1 &0 &0 & 0 \\
0 & 2\lambda_{32}^1 &1 &0 & 0 \\
0 & 2\lambda_{42}^1 &0 &\rho & 0 \\
0 & 0 &0 &0 & 0 \\
\end{bmatrix}.
$$

We implement in R.

```{r}
lambda <- t(matrix(
  c(-0.7,0.1,0.1,0.0,0.5,
    0,-0.5,0,0,0.5,
    0,0.1,-0.7,0.1,0.5,
    0,0.1,0,-0.6,0.5,
    0,0,0,0,0), ncol = 5,nrow = 5
))
lambda_1 <- t(matrix(
  c(0,0.05,0,0,0,
      0,0,0,0,0,
      0,0.1,0,0,0,
      0,0.05,0,0,0,
      0,0,0,0,0), ncol = 5,nrow = 5
))
lambda_0 <- lambda - lambda_1
R <- lambda_1 * B + diag(b)
```

</details>

<details>
<summary>**Solution (c).**</summary>

We compute the reserve by using the following

$$
\prod_t^T\left(\mathbf I+\begin{pmatrix}
\mathbf \Lambda-\Delta(r) & \mathbf R\\
\mathbf 0 & \mathbf \Lambda
\end{pmatrix}\ dx\right)=
\begin{pmatrix}
e^{-\int_t^Tr\ du}\mathbf P(t,T) & \mathbf V(t,T)\\
\mathbf 0 & \mathbf P(t,T)
\end{pmatrix}.
$$

We then compute 

$$
V(t,T)=e_1^\top \mathbf V(t,T)e.
$$

```{r}
r <- 0.04
big_matrix <- as.matrix(Matrix::bdiag(lambda-r*diag(rep(1,5),5),lambda))
big_matrix[1:5,6:10] <- R
van_loan <- Matrix::expm(big_matrix)
V <- c(1,0,0,0,0) %*% van_loan[1:5,6:10] %*% rep(1,5)
as.numeric(V)
```


</details>

<details>
<summary>**Solution (d).**</summary>

The equivalence premium is computed by

$$
\hat \rho = \left.-\frac{e_1^\top \mathbf V(t,T)e}{e_1^\top \mathbf V'(t,T)e}\right\vert_{\rho=0},
$$

with $\mathbf V'(t,T)$ defined by

$$
\prod_t^T\left(\mathbf I+\begin{pmatrix}
\mathbf \Lambda-\Delta(r) &\frac{\partial}{\partial \rho} \mathbf R(\rho)\\
\mathbf 0 & \mathbf \Lambda
\end{pmatrix}\ dx\right)=
\begin{pmatrix}
e^{-\int_t^Tr\ du}\mathbf P(t,T) & \mathbf V'(t,T)\\
\mathbf 0 & \mathbf P(t,T)
\end{pmatrix}.
$$

We compute

```{r}
rho <- 0
computeReserve <- function(rho,T){
  b <- c(rho, 1, 1, rho,0)
  R <- lambda_1 * B + diag(b)
  big_matrix <- as.matrix(Matrix::bdiag(lambda-r*diag(rep(1,5),5),lambda))
  big_matrix[1:5,6:10] <- R
  van_loan <- Matrix::expm(big_matrix*T)
  V <- c(1,0,0,0,0) %*% van_loan[1:5,6:10] %*% rep(1,5)
  return(as.numeric(V))
}
V <- computeReserve(rho,1)
R_prime <- diag(c(1,0,0,1,0))
big_matrix <- as.matrix(Matrix::bdiag(lambda-r*diag(rep(1,5),5),lambda))
big_matrix[1:5,6:10] <- R_prime
van_loan <- Matrix::expm(big_matrix)
V_prime <- c(1,0,0,0,0) %*% van_loan[1:5,6:10] %*% rep(1,5)
V_prime <- as.numeric(V_prime)

rho_hat <- - V / V_prime
rho_hat
```


Let us see if the reserve is zero as expected.

```{r}
computeReserve(rho_hat,1)
```


</details>

<details>
<summary>**Solution (e).**</summary>

We have that

$$
\mathbf C_r^{(k)}=\frac{1}{k!}\mathbf\Lambda^1\bullet \mathbf B^{\bullet k}
$$

and
\begin{align*}
\prod_t^T&\left(\mathbf I+\begin{pmatrix}
\mathbf \Lambda-k\Delta(r) & \mathbf R & \mathbf C_r^{(2)} & \cdots & \mathbf C_r^{(k)}\\
\mathbf 0 & \mathbf \Lambda - (k-1)\Delta(r) & \mathbf R & \cdots & \mathbf C_r^{(k-1)}\\
\vdots & \vdots & \vdots & \ddots & \vdots \\
\mathbf 0 & \mathbf 0 & \mathbf 0 & \cdots & \mathbf \Lambda
\end{pmatrix}\ dx\right)\\
&=
\begin{pmatrix}
* & * & * & \cdots & \mathbf V_r^{(k)}(t,T)\\
* & * & * & \cdots & \mathbf V_r^{(k-1)}(t,T)\\
\vdots & \vdots & \vdots & \ddots & \vdots \\
* & * &* & \cdots &  \mathbf P(t,T)
\end{pmatrix}.
\end{align*}
Where

$$
\mathbf V_r^{(k)}(t,T)=\frac{1}{k!}\mathbf V^{(k)}(t,T).
$$

We therefore implement a Toeplizs block matrix function that for matrices $\mathbf M_1,...,\mathbf M_k$ generates the matrix

$$
\mathbf M=\begin{pmatrix}
\mathbf M_1 & \mathbf M_2 & \mathbf M_3 &\cdots & \mathbf M_k\\
\mathbf 0 & \mathbf M_1 & \mathbf M_2 & \cdots & \mathbf M_{k-1}\\
\mathbf 0 & \mathbf 0 & \mathbf M_1 & \cdots & \mathbf M_{k-2}\\
\vdots & \vdots & \vdots & \ddots & \vdots\\
\mathbf 0 & \mathbf 0 & \mathbf 0 & \cdots & \mathbf M_1
\end{pmatrix}
$$

This is done in the `bToeplitz`

```{r,message=FALSE}
bToeplitz <- function(m) {
  n <- length(m) #number of diagonals
  p <- unique(as.numeric(sapply(1:n, function(i) dim(m[[i]])))) #dimensions
  if (length(p) > 1) {
    print("Aborting: Dimensions is not equal!")
  } else {
    require(Matrix)
    result <- matrix(rep(0,p**2*n**2),ncol = n*p)
    for (i in 1:n) {
      m_diag <- m[[i]]
      increment <- as.matrix(Matrix::bdiag(
        lapply(1:(n-i+1), function(j){
          m_diag
        })
      ))
      result[1:((n-i+1)*p),(p*(i-1)+1):(n*p)] <- result[1:((n-i+1)*p),(p*(i-1)+1):(n*p)]+ increment
    }
    return(result)
  }
}
#An example
bToeplitz(list(lambda,R))
```

We can then compute the second moment.

```{r}
b <- c(rho_hat,1,1,rho_hat,0)
big_matrix <- bToeplitz(
  list(lambda,
       lambda_1*B + diag(b),
       (1/factorial(2))*lambda_1*B**2)
) - as.matrix(Matrix::bdiag(2*r*diag(1,5),
                            r*diag(1,5),
                            0*diag(1,5)))
van_loan <- expm::expm(big_matrix)
#Second moment
V2 <- c(1,0,0,0,0) %*% (factorial(2)*van_loan[1:5,11:15] %*% rep(1,5))
V2
V <- c(1,0,0,0,0) %*% van_loan[6:10,11:15] %*% rep(1,5)
V
```

We can then see that the variance is `r round(V2-V**2,2)`.

</details>
